RadLib: a radiative heat transfer model library for CFD

### 2021-04-08 submissions/CPC_response_to_reviewers.md

**TO DO: consult Josh about appropriate tests for RadLib**

5. CMake build system:
    1.	Consider installing a “RadLib.cmake” file for downstream usage by CMake-based projects. This helps downstream build systems configure for RadLib usage (setting include paths, etc.).
    1.	probably shouldn’t specify optimize or debug flags directly, as CMake will provide appropriate values on most platforms.
    1.	the default build type isn’t defaulting to Release for me; it remains blank.
    1.	I suggest not using verbose makefiles by default.
    1.	Using GLOB to install the headers is not the appropriate CMake approach.
    1.	the data files are not copied into the examples build directory, meaning that the example executables do not run.
    1.	The run-examples.sh script is not helpful as it is not installed into the build directory where the executables are produced and doesn’t have path information to actually run the executables.
    1.	Using “doxygen” as a target is not ideal since it conflicts with the “doxygen” executable name. It still works fine though...

5. It appears that line-by-line data from HITEMP/HITRAN files cannot be loaded into RadLib. That’s an unfortunate limitation, but not a deal-breaker.

	**At present, RadLib's models cannot accommodate such data input, but this is a limitation of the models themselves, not the software. The appropriate line-by-line data is included within the RadLib package, which the authors consider a convenience rather than a hindrance. The ability to load database files directly into RadLib could be useful to researchers looking to extend RadLib's model library to include modern spectral models, however. The authors appreciate the suggestion and will consider it for future development.**

@@

	**The authors purposely chose the TNF workshop data for calculating Planck Mean absorption coefficients because it is one of the most common and widely used radiation models applied to turbulent combustion simulations, despite its age and limitations. The paper includes a brief discussion on the limited applicability of this model, especially for combustion simulations, and demonstrates its deficiencies in the example cases. It is true that this model is neither current nor particularly accurate, and the authors are strongly considering adding an updated Planck Mean model using the most current available data.**

	**TO DO: add discussion to address the age of the model and why we chose to include it anyway**

5) On page 13, the authors present a paragraph describing Figs. 2 and 3 - the comparison of various radiation property models present in RadLib in one-dimensional configurations. The comparisons are presented simply as observations without any explanation or reasoning why the models behave the way as seen in Fig. 2 and 3. If I understand correctly, the focus of the work presented here is not the comparison of the accuracy of the three models but to showcase the library. Therefore, these comparisons - specifically without any explanation of the results - may not be essential to the current manuscript, and can be shortened. These results do serve as descriptions of example cases provided in RadLib. On the same page, the authors say that "While ommitted [should be omitted] here for brevity, the implemented WSGG and RCSLW models give essentially identical results to those presented in [11, 14] such that these examples also serve as a validation of the implementation of the
models." The authors should avoid qualitative description such as "essentially identical" in the context of validation and present a quantitative (e.g., % error) metric wherever possible. A strong validation is important for programs like this.

**TO DO: decide on whether or not to keep example plots in the paper**

@@

6) The computational cost tables for two cases show a significantly different trend, which the authors attribute to "additional overhead required in performing the calculations." Can the authors be more specific as to what these overheads may be? They say that they have neglected the initialization and input/output costs. Also, do these CPU times presented include the cost for the ray-tracing solver as well? Are the grids kept the same for all configurations?

**TO DO**

7) In line 53 on page 4, the authors claim to have CH4 property in RadLib but none of the validation cases contain CH4. Furthermore, for RCSLW, the refer to Pearson's work as a source for data of CO2, H2O, and CO without indicating the source data for CH4.

	**The Plank Mean absorption coefficients model sourced from the TNF workshop and associated literature [3,5] includes CH4, so it can be done, but none of the examples include it.**

**TO DO: ask DOL how to handle this**

@@

MINOR:

1) It should be clarified that RadLib only provides radiation properties for non-scattering media.

	**This clarification has been made explicitly in Section 2.**

	**TO DO: make sure this is mentioned in other relevant places**

@@

2) In the program summary the description of the nature of the program is somewhat confusing from a grammatical construction standpoint. Particularly, the following part, "This presents a problem for practitioners who wish to use/implement radiation models for which [refers to what?] such models consitute [constitute?] only a small but important part of a larger simulation.  Turbulent combustion simulations are one such example [example of what?].  Often, rudimentary assumptions [in what?] are made and this can negatively impact results."

**TO DO: revise program summary**

3) In Page 2, "In some cases, radiative gains or losses are small compared to other energy sources or heat transfer modes and radiation can be safely neglected, ... difficult to simulate accurately without a robust radiation model." Please provide some references to support these arguments.

	**References have been added throughout to bolster weakly supported arguments. Note that the quoted statement and related discussion has been moved to Section 5 due to unrelated revisions.**

### 2021-04-12 manuscript/manuscript.tex
Revised and rewrote PM model section

For an absorbing, emitting, and non-scattering medium, the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$. Solving the RTE requires information about the radiative properties of the materials and media in the system. Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model. Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules, reflecting varying degrees of dependence on local system properties like temperature, pressure, and gas composition. Radiation property models, including those implemented in RadLib, typically use high-resolution spectral databases like HITRAN and HITEMP \cite{Rothman_2010} to calculate spectral absorption coefficients, reducing or simplifying the huge amount of spectral data using correlations and modeling assumptions \cite{Zhang_2002b}.

@@

With RadLib, we aim to provide researchers with a convenient, easy-to-use access point for radiative property models that can be interfaced with various RTE solution approaches suitable for CFD applications. RadLib is a modular C++ library with an optional Python interface. Currently, RadLib only calculates radiation absorption coefficients and weighting factors for use in global radiation models, but expansion to other radiative properties and model types is possible. Three models---the Planck Mean absorption coefficient (PM) model, the weighted sum of gray gases (WSGG) model, and the rank correlated spectral line weighted-sum-of-gray-gases (RCSLW) model---are fully implemented and validated. All models use a common interface to facilitate researchers' access to and use of radiation modeling tools suitable for combustion CFD applications.

@@

At present, RadLib includes three radiation property models of varying complexity and accuracy.

 to calculate the radiation absorption coefficients $\kappa_j$ and their weight factors $a_j$ for each gas $j$. The absorption coefficients are typically calculated using correlations relating them to local properties such as temperature, pressure, and composition. Correlations come from curve fits to high-resolution radiation property databases. At present, RadLib considers up to four gas species (H$_2$O, CO, CO$_2$, and CH$_4$) and soot. Note that the property models implemented in RadLib apply only to non-scattering media.

While solving the RTE is not the focus of this software, RadLib's example cases do employ a simple ray tracing solver to calculate the radiative heat flux and volumetric heat source profiles between two parallel planes.

Absorption coefficients and other radiative property data may take on different forms depending on the RTE solution approach, but can generally be classified as either spectral (wavenumber-dependent) or global (integrated over wavenumber). Spectral absorption coefficients can be applied to the spectral RTE (\ref{e:RTE_general}), while global absorption coefficients are used in the spectrally-integrated RTE (\ref{e:RTE_wsgg}).

@@

RadLib's simplest model uses Planck Mean absorption coefficients calculated from the correlations given on the TNF Workshop site \citep{Smith_2003}. The temperature-dependent correlations are based on the RADCAL model \citep{Grosshandler_1993}. The TNF radiation model is also documented in \citep{Barlow_2001}.

Planck Mean absorption coefficients are commonly used to model radiation, especially in combustion systems, because the model is relatively easy to implement, computationally inexpensive, and reasonably accurate in some cases. The Planck Mean model is most appropriate under optically thin conditions in which low radiative transfer relative to other heat sources, including but not limited to chemical heat release and buoyancy-driven convection in combustion systems, indicates that very little radiative reabsorption is occuring. In cases where the optically thin assumption applies, such as simple hydrogen jet flames, this model and the associated Planck Mean absorption coefficients can produce accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, which includes most practical combustion scenarios of interest to engineers, the Planck Mean model does not produce accurate results. For example, it significantly overpredicts radiative losses from the TNF library's CH$_4$ flames \cite{Frank_2000,Zhu_2002,Coelho_2002}. The examples included with the code and illustrated in Section \ref{s:Examples} demonstrate the deficiencies of the Planck Mean model in comparison to the others included with RadLib.

>>>>

For an absorbing, emitting, and non-scattering medium, the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$. Solving the RTE requires information about the radiative properties of the materials and media in the system. Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model. Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules, reflecting varying degrees of dependence on local system properties like temperature, pressure, and gas composition. Radiation property models, including those implemented in RadLib, typically use high-resolution spectral databases like HITRAN and HITEMP \cite{Rothman_2010} to calculate spectral absorption coefficients, reducing or simplifying the huge amount of spectral data using correlations and modeling assumptions.

@@

With RadLib, we aim to provide researchers with a convenient, easy-to-use access point for radiative property models that can be interfaced with various RTE solution approaches suitable for CFD applications. RadLib is a modular C++ library with an optional Python interface. Currently, RadLib only calculates radiation absorption coefficients and weighting factors for use in global radiation models, but expansion to other radiative properties and model types is possible. Three models are fully implemented and validated. All models use a common interface to facilitate researchers' access to and use of radiation modeling tools suitable for combustion CFD applications.

@@

At present, RadLib includes three radiation property models of varying complexity: Planck Mean absorption coefficients (assuming optically thin media), a weighted sum of gray gases (WSGG) model, and the rank-correlated spectral line weighted-sum-of-gray-gases (RCSLW) model. Each of these models was specifically chosen for its notable qualities, attributes such popularity, ease of implementation, or accuracy; their individual strengths, weaknesses, and limitations are discussed in the sections that follow.

Absorption coefficients and other radiative property data may take on different forms depending on the RTE solution approach, but can generally be classified as either spectral (wavenumber-dependent) or global (integrated over wavenumber). Spectral absorption coefficients can be applied to the spectral RTE (Eq.~\ref{e:RTE_general}), while global absorption coefficients are used in the spectrally-integrated RTE (Eq.~\ref{e:RTE_wsgg}). The type of absorption coefficient generated by each model and potential RTE integration strategies will also be discussed in the following sections. While solving the RTE is not the focus of this software, RadLib's example cases do employ a simple ray tracing solver to calculate the radiative heat flux and volumetric heat source profiles between two parallel planes. Refer to Section \ref{s:Examples} for details.

Radiation absorption coefficients are typically calculated by correlations developed using high-resolution databases like RADCAL, HITEMP, and HITRAN \citep{Grosshandler_1993,Rothman_2010} that relate spectral lines to properties such as temperature, pressure, and gas composition \citep{Zhang_2002b}. At present, RadLib considers up to four gas species (H$_2$O, CO, CO$_2$, and CH$_4$) and soot. Note that the property models implemented in RadLib apply only to non-scattering media.

@@

RadLib's simplest radiation property model generates Planck Mean absorption coefficients under the assumption of optically thin media, from here on referred to as the Planck Mean (PM) model.
The optically thin approximation assumes that there is little to no reabsorption of radiative energy in the system, or, in other words, that the system only emits radiation and does not absorb it. In combustion systems, optical thinness is often justified in terms of other flame properties such as low radiative heat transfer relative to chemical heat release or convective heat transfer (also used to justify neglecting radiation altogether in some cases). Assuming optical thinness, however, has the distinct advantage of simplifying the radiative source term (here denoted $S_{rad}$) of the energy conservation equation to 
\begin{equation}
	S_{rad}=-4\kappa \sigma T^4,
\end{equation}
where $\kappa$ is the radiation absorption coefficient, $\sigma$ is the Stefan-Boltzmann constant, and $T$ is temperature. Using an optically thin approximation means that no RTE solution is required when calculating radiative source terms.

In combustion systems, this simplification is convenient because it makes radiation modeling easy to implement and computationally inexpensive. The vast majority of computational combustion research that considers radiation employs an optically thin approximation despite its limited applicability to combustion processes. Assuming optical thinness can predict the radiative behavior of small laboratory flames reasonably well, but in larger flame or industrial systems tends to overpredict radiative heat losses by 50\% or more and underpredicts flame temperatures by over 100$^{\circ}$C \citep{Modest_2016}.

The specific correlations for the Planck Mean absorption coefficients generated by RadLib come from the Turbulent Nonpremixed Flame (TNF) Workshop, an ongoing collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. These particular correlations were chosen for RadLib's PM model because of the popularity and widespread use of the TNF Workshop radiation model in combustion research despite its notable limitations. Its temperature-dependent correlations are based on the RADCAL database \citep{Grosshandler_1993}, which has been largely superseded by more comprehensive and accurate databases, particularly HITRAN and HITEMP \citep{Rothman_2010}. TNF Workshop literature shows that in cases where the optically thin approximation accurately describes the system, such as in simple hydrogen jet flames, the PM model produces accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, however, including most practical combustion scenarios of interest to scientists and engineers, the PM model does not produce accurate results, overpredicting radiative heat loss and underpredicting flame temperature by 100$^{\circ}$ or more \cite{Frank_2000,Zhu_2002,Coelho_2002}.

Because of these limitations, the authors do not consider the PM model adequate for most combustion simulation cases and strongly recommend that RadLib users carefully consider its implications before using the PM model. It is implemented in RadLib primarily for comparison and evaluation against other radiation property models because of its popularity and widespread use. The examples included with the RadLib code and shown in Section \ref{s:Examples} demonstrate the deficiencies of the Planck Mean model in comparison to other models included with RadLib.
